Customizing Software by Direct Manipulation of Tabular Data

### 2020-05-22 paper.tex
first final

Sifter \citep{huynh2006} enables end users to sort and filter lists of
data obtained by web scraping, much like Wildcard's sorting features.
The main difference between the systems is that table-driven
customization has many other use cases besides sorting and filtering.
Also, Sifter involves end users in a semi-automated data extraction
process, rather than having programmers create wrappers. This provides
coverage of more websites, but at the expense of complicating the end
user experience. We might explore such techniques in Wildcard in the
future, but we think that it's valuable for end users to have a
customization experience decoupled from the challenge of web scraping
the underlying data. Sifter implements scraping across multiple pages, a
valuable feature for sorting and filtering that isn't present in
Wildcard.

Thresher \citep{hogue2005} helps end users create wrappers that map
website content to Semantic Web schemas, making customizations available
using the schema information. Wildcard shares the general idea of
wrappers with Thresher, but maps to a single generic data type, rather
than more specific schemas, increasing the range of websites and data
that Wildcard is able to support.

There are many software customization tools that offer simplified forms
of programming for end users. Chickenfoot \citep{bolin2005} and
Coscripter \citep{leshed2008} offer user friendly syntax for writing web
automation scripts; Applescript \citep{cook2007} has a similar goal for
desktop customization. There are visual programming environments for
customization that don't involve writing any text:
\href{https://support.apple.com/guide/automator/welcome/mac}{Automator}
for Mac and
\href{https://apps.apple.com/us/app/shortcuts/id915249334}{Shortcuts}
for iOS are modern options for customizing Apple products, and
\href{https://zapier.com/}{Zapier} enables users to connect different
web applications together visually. As mentioned previously, these tools
all require writing imperative programs in some form, in contrast to the
more declarative and direct approach of table-driven customization.

@@

Another relevant area involves direct manipulation interfaces for
interacting with structured data. We take inspiration from these tools
in our work, but apply them in a different domain: customizing existing
software applications, rather than interacting with databases or
constructing software from scratch.

The most closely related work is in systems that offer spreadsheet-like
querying of relational data, as proposed by Liu and Jagadish
\citep{liu2009}. SIEUFERD by Bakke and Karger \citep{bakke2016} is an
example; their paper presents a survey of other similar tools. Our work
is particularly influenced by the authors' observation that direct
manipulation requires that the user manipulate the results of a database
query rather than the query itself, and that the user must see
intermediate results at every step of constructing the query. SIEUFERD's
interface supports a far wider range of queries than Wildcard, but the
basic ideas of the user interface are similar.

\href{https://airtable.com/}{Airtable} is another example of a modern
commercial product that offers spreadsheet-like interaction with a
relational database.

Our work is also inspired by the many projects that have explored using
spreadsheets as a foundation for building software applications,
including Object Spreadsheets \citep{mccutchen2016}, Quilt
\citep{benson2014}, Gneiss \citep{chang2014}, Marmite \citep{wong2007},
\href{https://www.glideapps.com/}{Glide}. These projects share the idea
that a spreadsheet is a convenient interface for editing the data
underlying a GUI application. We share that idea, but apply it to
software customization, rather than building software from scratch.

Another system in this space is ScrAPIr, by Alrashed et
al.~\citep{alrashed2020}, which enables end users to access backend web
APIs without programming. ScrAPIr shares our high level goal of end user
empowerment, as well as the idea of wrappers, by creating a shared
library of wrappers around existing web APIs. Unlike Wildcard, however,
ScrAPIr targets explicit APIs exposed by developers. It also focused on
backend services and doesn't aim to extend the frontend interfaces of
web applications.

>>>>

Sifter \citep{huynh2006} enables end users to sort and filter lists of
data obtained by web scraping, much like Wildcard's sorting features.
The main difference between the systems is that table-driven
customization has many other use cases besides sorting and filtering.
Also, Sifter involves end users in a semi-automated data extraction
process, rather than having programmers create wrappers. This provides
coverage of more websites, but at the expense of complicating the end
user experience. We could integrate end user scraping techniques in
Wildcard in the future, but we believe that when possible it's valuable
for end users to have a customization experience decoupled from the
challenge of web scraping the underlying data. Sifter also implements
scraping across multiple pages, a valuable feature for sorting and
filtering that isn't present in Wildcard.

Thresher \citep{hogue2005} helps end users create wrappers that map
website content to Semantic Web schemas like ``Movie'' or ``Director,''
and augments websites with new functionality by exploiting that schema
information. Wildcard shares the general idea of wrappers, but maps to a
generic table data type rather than more specific schemas, increasing
the range of supported data and allowing for a simpler mapping process.

There are many software customization tools that offer simplified forms
of programming for end users. Chickenfoot \citep{bolin2005} and
Coscripter \citep{leshed2008} offer user friendly syntax for writing web
automation scripts; Applescript \citep{cook2007} has a similar goal for
desktop customization. There are visual programming environments for
customization that don't involve writing any text:
\href{https://support.apple.com/guide/automator/welcome/mac}{Automator}
for Mac and
\href{https://apps.apple.com/us/app/shortcuts/id915249334}{Shortcuts}
for iOS are modern options for customizing Apple products, and
\href{https://zapier.com/}{Zapier} enables users to connect different
web applications together visually. As mentioned previously, these tools
all require writing imperative programs, in contrast to the more
declarative and direct approach of table-driven customization.

@@

Another relevant area involves spreadsheets and visual query systems. We
take inspiration from these tools in our work, but apply them in a
different domain: customizing existing software applications, rather
than interacting with databases or constructing software from scratch.

The most closely related work is in systems that offer spreadsheet-like
querying of relational data. SIEUFERD by Bakke and Karger
\citep{bakke2016} is one such recent system, and their paper presents a
survey of many other similar tools. Our work is particularly influenced
by the authors' observation that direct manipulation requires that the
user manipulate the results of a database query rather than the query
itself, and that the user must see intermediate results at every step of
constructing the query. SIEUFERD's interface supports a far wider range
of queries than Wildcard, but the fundamental principles of the user
interface are similar. \href{https://airtable.com/}{Airtable} is another
example of a modern commercial product that offers spreadsheet-like
interaction with a relational database.

Our work is also inspired by the many projects that have explored using
spreadsheets as a foundation for building software applications,
including Object Spreadsheets \citep{mccutchen2016}, Quilt
\citep{benson2014}, Gneiss \citep{chang2014}, Marmite \citep{wong2007},
\href{https://www.glideapps.com/}{Glide}. We share the core idea of
connecting a spreadsheet view to a GUI, but apply it to software
customization, rather than building software from scratch.

Another related system is ScrAPIr, by Alrashed et
al.~\citep{alrashed2020}, which enables end users to access backend web
APIs without programming. ScrAPIr shares our high level goal of end user
empowerment, as well as the idea of wrappers, by creating a shared
library of wrappers around existing web APIs. Unlike Wildcard, however,
ScrAPIr targets explicit APIs exposed by developers. It also focused on
backend services and doesn't aim to extend the frontend interfaces of
web applications.
